---
title: "R Notebook"
output: html_notebook
---

---
title: "climate change"
author: "Carlos aybar"
output: html_notebook

---

### install tidyverse package
```{r}
install.packages("tidyverse")
install.packages("corrplot")
install.packages("Metrics")

```


### load all the necessary libraries
```{r}
library(tidyverse)
library(ggplot2)
library(corrplot)
library(Metrics)
```

### importing and structuring the data
```{r}
climate_change <- read.csv("climate_change.csv")

structure(climate_change)
```

# linear regression for carbon dioxide over the years
Given 1 independent variable, predict the dependent variable.
Given the goal difference, let's predict the team's final position.
```{r}
model <- lm(carbon_dioxide ~ Year, data = climate_change)
#predicting the amount of co2 in the atmosphere by 2050
test <- data.frame(Year = c(2050))

# Predict the position with the model and test data
predictedCO2 <- predict(model, test)
cat("the amount of co2 in the atmosphere by 2050 will be: ", predictedCO2)

```


# first correlation plot

```{r}
plot(climate_change$carbon_dioxide, climate_change$Year, xlab = "Carbon Dioxide (ppmv)", ylab= "Year", main = "Carbon dioxide production over the years", col = "red")
lines(lowess(climate_change$carbon_dioxide, climate_change$Year), col = "blue")
```


# lm model for temperature change based on methane production
Given 1 independent variable, predict the dependent variable.
Given the goal difference, let's predict the team's final position.
```{r}
model <- lm(temp_change ~ methane, data = climate_change)
#predicting the change in temperature when methane level is 2000
test <- data.frame(methane = c(2000))

# Predict the temperature change in celsius with the model and test data
predictedTemperature <- predict(model, test)
cat("the change in temperature when there is a methane production of 2000 ppmv is: ", predictedTemperature, "°C")

```
# Correlation: Pairs Plot


```{r}

plot(climate_change$methane, climate_change$temp_change, xlab = "Methane(ppmv)", ylab= "Temperature Change(°C)", main = "Effect methane on temperature change", col = "blue")
lines(lowess(climate_change$methane, climate_change$temp_change), col = "red")

```
# lm model for the effect of temperature change on El Niño
Given 1 independent variable, predict the dependent variable.
Given the goal difference, let's predict the team's final position.
```{r}
model <- lm(MEI ~ temp_change, data = climate_change)
#predicting the change in MEI when temperature chagnes by 1 degree celcius
test <- data.frame(temp_change = c(1))

# Predict the temperature change in celsius with the model and test data
predictedMEI <- predict(model, test)
cat("If the temperature changes by 1°C the IME will change by: ", predictedMEI, "ppmv")


```

#visualization

```{r}
plot(climate_change$MEI, climate_change$temp_change, xlab = "multivariate El Niño Index", ylab= "Temperature Change", main = "Effect of temperature change on the intensity of an El Niño Southern Oscillation", col = "red")
lines(lowess(climate_change$MEI, climate_change$temp_change), col = "black")
```



# lm model for the effect of aerosols on solar irradiance
Given 1 independent variable, predict the dependent variable.
Given the goal difference, let's predict the team's final position.
```{r}
model <- lm(totalsolar_irradiance ~ Aerosols, data = climate_change)
#predicting the change solar irradiance when aerosols are 0.1863
test <- data.frame(Aerosols = c(0.1863))

# Predict the temperature change in celsius with the model and test data
predictedTSI <- predict(model, test)
cat("0.1863nm aerosols will lead to : ", predictedTSI, "solar irradiance per unit area")

```

#visualization

```{r}
plot(climate_change$Aerosols, climate_change$totalsolar_irradiance, xlab = "Aerosols", ylab= "total solar irradiance", main = "Effect aerosols on total irradiance", col = "orange")
lines(lowess(climate_change$Aerosols, climate_change$totalsolar_irradiance), col = "green")

```




# lm model to predict N2O production each month
Given 1 independent variable, predict the dependent variable.
Given the goal difference, let's predict the team's final position.
```{r}
model <- lm(nitrous_oxide ~ Year, data = climate_change)
#predicting the change N2O by 2030
test <- data.frame(Year = c(2030))

# Predict the temperature change in celsius with the model and test data
predictedN2O <- predict(model, test)
cat("the amount of nitrous oxide in the air by 2030 will be: ", predictedN2O, "ppmv")

```

#visualization

```{r}
plot(climate_change$nitrous_oxide, climate_change$Year, xlab = "nitrous oxide", ylab= "Year", main = "Nitrous oxide production since 1980 year", col = "green")
lines(lowess(climate_change$nitrous_oxide, climate_change$Year), col = "blue")
```
